QPLIB: A Library of Quadratic Programming Instances

### 2017-01-20 QPLIB-1.0.tex
update/add todo

Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last 60 years. At least some of the rationale for this interest is likely due to the fact that they are the ``least-nonlinear nonlinear problems''. Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP). This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others ({\bf citations?}). Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved. Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.

On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP. Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}. Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with either quadratic constraints, ot quadratic objective function, or both. Also, in general any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions. Hence, (MI)QP is, in some sense, roughly as expressive as the whole of (MI)NLP. Of course this is, in principle, true for (MI)LP as well, but at the cost of much larger and much more complex formulations ({\bf citations?}). Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of (MI)LP and the higher expressive power, but much higher computational cost, of (MI)NLP.

>>>>

Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last 60 years. At least some of the rationale for this interest is likely due to the fact that they are the ``least-nonlinear nonlinear problems''. Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP). This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \todo[inline]{citations? http://www.optimization-online.org/DB\_HTML/2009/07/2338.html ?}. Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved. \todo[inline]{If only the objective is quadratic, i.e., what one usually would denote by a ``nonconvex QP'', then one can ``just'' enumerate local optima. Nitpicking?} Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.

On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP. Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}. Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with either quadratic constraints, or quadratic objective function, or both. Also, in general any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions. Hence, (MI)QP is, in some sense, roughly as expressive as the whole of (MI)NLP. Of course this is, in principle, true for (MI)LP as well, but at the cost of much larger and much more complex formulations ({\bf citations?}). Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of (MI)LP and the higher expressive power, but much higher computational cost, of (MI)NLP.

### 2017-01-20 QPLIB.tex
update/add todo

\begin{abstract}
 This paper describes a new instance library for Quadratic Programming (QP).
 QP is a very ``varied'' class, comprising sub-classes of problems ranging from
 trivial to undecidable. Solution methods for QP are very diverse, ranging from
 entirely combinatorial ones to completely continuous ones, including many for
 which both aspects are fundamental. Selecting a set of instances of QP that is
 at the same time not overwhelmingly numerous and sufficiently challenging for the
 many different interested communities is therefore important. We propose a simple
 taxonomy for QP instances that leads to a systematic problem selection mechanism.
 We then briefly survey the field of QP, giving an overview of theory, methods and
 solvers. Finally we describe how the library was put together, and detail its final
 contents.
\end{abstract}

>>>>

\begin{abstract}
 This paper describes a new instance library for Quadratic Programming (QP).\todo[inline]{As we don't use the usual definition of QP, we should already say in the abstract what we understand under a QP.}
 QP is a very ``varied'' class, comprising sub-classes of problems ranging from
 trivial to undecidable. Solution methods for QP are very diverse, ranging from
 entirely combinatorial ones to completely continuous ones, including many for
 which both aspects are fundamental. Selecting a set of instances of QP that is
 at the same time not overwhelmingly numerous and sufficiently challenging for the
 many different interested communities is therefore important. We propose a simple
 taxonomy for QP instances that leads to a systematic problem selection mechanism.
 We then briefly survey the field of QP, giving an overview of theory, methods and
 solvers. Finally we describe how the library was put together, and detail its final
 contents.
\end{abstract}
